Lemma 1

### 2021-03-15 notes/jessie-russian-lemma.tex
Updating Russian lemma with refs and cleaning

\begin{document}
\maketitle
	\jessie{Meeting notes Mar 1, 2021: - Careful about pulling out expectation in bullet 2; when is inner product of RV = expectation of inner product? - Point out contradicting the second line would yield contradiction of subgradient inequality - Go slowly through the definition of measurable}
\paragraph{Setting} 
$\effdom(L) = \reals^d$, and $\E_p L(u,Y) < \infty$ for all $u \in \reals^d$ and $p \in \P$ \jessie{$y \in \Y$?}

@@

We define $\E_p \partial L(u,Y) = \{\E_p V: V(y) \in \partial L(u,y) p\text{-almost surely}, V \text{ measurable}\}$.
Moreover, let $\V_{u} = \{V : \Y \to \reals^d : V \text{ measurable}, V(\omega) \in \partial L(u,\omega) \forall \omega \in \Omega\}$.
Observe the set $\Z_u(p) := \{\E_p V(Y): V \in \V_u\} \subseteq \E_p \partial L(u,Y)$.

@@

Fix $z$.
Consider the set-valued function $Q_z : \Omega \to 2^{\reals^d}$ defined by $Q_z(\omega) := \{x \in \partial L(u,\omega) : \inprod{z}{x} = L'(u,\omega;z)\}$.
Now, we want to take a measurable function $V \in \V_u$ that is a selection of $Q_z$ satisfying the following conditions:
\begin{enumerate}
	\item $V \in \V_u$.
	\item $\E_p V \in \E_p \partial L(u,Y)$.
	\item $\inprod{z}{\E_p V} = \E_p L'(u,Y,z)$
\end{enumerate} 
If these are true, then we can take $v := \E_p V$ and observe the result.

@@

1.  
If $Q_z$ is measurable and closed-valued, then by \jessie{Castaing, also Rockafellar VA 14.6}, we have a measurable selection $V \in \V_u$ such that $\inprod{z}{V(\omega)} = L'(u,\omega; z)$ for all $\omega \in \Omega$.

@@

In order to show that $Q_z$ is $B$-measurable, it is equivalent to show $gr(Q_z)$ is $B \otimes \B(\reals^d)$-measurable. \jessie{Kinda hand-wavy} \jessie{ASSUMPTION: $B$ is a complete $\sigma$-algebra with respect to $p$.}
Now, consider $gr(Q_z) = gr(Q) \cap gr(R)$, where $Q: \omega \mapsto \partial L(u,\omega)$ and $R : \omega \mapsto \{x \mid \inprod{z}{x} = L'(u,\omega;z)\}$.
As $Q$ is measurable \jessie{Rockafellar: Measurable Dependence, Cor 4.6}, so is $gr(Q)$.
Moreover, $gr(R)$ is measurable as it is the graph of a linear multifunction, hence measurable.
Now we can take $gr(Q_z) = gr(Q) \cap gr(R)$, which is measurable by closure under intersections, and have the result.

\bigskip
\hrulefill

\jessiet{If $R$ was compact}
As measurable functions are closed under (countable) intersections, we can simply show $Q_z$ intersection of measurable functions; namely, since we have $Q_z(\omega) = Q: \omega \mapsto \partial L(u,\omega) \cap R : \omega \mapsto \{x \in \reals^d : \inprod{z}{x} = L'(u,\omega;z)\}$.
\jessiet{Look at strong measurability or see if there's a result of closure under intersections}
\jessie{If $R$ is compact, then we are done by Castaing Valadier Prop iii.4}

Moreover, the function $Q$ is measurable by \jessie{Rockafellar: Measurable Dependence, Corollary 4.6}.
Moreover, $R$ is measurable as $L'$ is measurable and $\inprod{z}{\cdot}$ is continuous in its second argument, hence Borel.

As $Q_z$ is the intersection of these two functions, we conclude $Q_z$ is measurable by \jessie{Castaing Valadier Prop iii.4} and closed-valued, and therefore has a measurable selection $V \in \V_u$ such that $\inprod{z}{V(\omega)} = L'(u,\omega;z)$ for all $\omega \in \Omega$.

\hrulefill

@@

3.
We want to show $\inprod{z}{v} = \E_p L'(u,Y,z)$.
This follows from linearity of expectation and construction of $V$; to see this, consider
\begin{align*}
\inprod{z}{v} &= \inprod{z}{\E_p V} \\
 &= \E_p \inprod{z}{V} \jessie{\text{Use finiteness of expectation of } V}
 \end{align*}
 The expectation can be pulled out of the inner product because $V$ is finite for all $\omega \in \Omega$ as it is contained in $\partial L(u,\omega)$, which is compact as $L$ is finite and convex for all $u \in \reals^d$.
 \begin{align*}
 \E_p \inprod{z}{V} &= \int \inprod{z}{V(\omega)} dp\omega \\
 &= \int L'(u,\omega; z) dp\omega \\
 &= \E_p L'(u,Y;z)
\end{align*}

@@

>>>>

\begin{document}
\maketitle
\paragraph{Setting} 
$\effdom(L) = \reals^d$, and $\E_p L(u,Y) < \infty$ for all $u \in \reals^d$ and $p \in \P$ \jessie{$y \in \Y$?}

@@

We define $\E_p \partial L(u,Y) = \{\E_p V: V(y) \in \partial L(u,y) p\text{-almost surely}, V \text{ measurable}\}$.
Moreover, let $\V_{u} = \{V : \Y \to \reals^d : V \text{ measurable}, V(\omega) \in \partial L(u,\omega) \forall \omega \in \Omega\}$.
Observe the set $\Z_u(p) := \{\E_p V(Y): V \in \V_u\} \subseteq \E_p \partial L(u,Y)$.
Consider the measurable space $(\Omega, B)$, and assume $B$ is a complete $\sigma$-algebra with respect to $p$. \jessiet{So we get equivalence of measurability of the function and the graph}

@@

Fix $z$.
Consider the set-valued function $Q_z : \Omega \to 2^{\reals^d}$ defined by $Q_z(\omega) := \{x \in \partial L(u,\omega) : \inprod{z}{x} = L'(u,\omega;z)\}$.
Now, we want to take a measurable selection $V$ of $Q_z$ satisfying the following conditions:
\begin{enumerate}
	\item $V \in \V_u$.
	\item $\E_p V \in \E_p \partial L(u,Y)$.
	\item $\inprod{z}{\E_p V} = \E_p L'(u,Y,z)$
\end{enumerate} 
If these are true, then we can take $v := \E_p V$ and observe the result.

@@

1.  
If $Q_z$ is measurable and closed-valued, then by~\cite[Theorem III.6]{castaing2006convex} \jessiet{Also Rockafellar VA 14.6}, we have a measurable selection $V \in \V_u$ such that $\inprod{z}{V(\omega)} = L'(u,\omega; z)$ for all $\omega \in \Omega$.

@@

In order to show that $Q_z$ is $B$-measurable, it is equivalent to show $gr(Q_z)$ is $B \otimes \B(\reals^d)$-measurable since $B$ is a complete $\sigma$-algebra with respect to $p$~\cite[Chapter III]{castaing2006convex}. 
Now, consider $gr(Q_z) = gr(Q) \cap gr(R)$, where $Q: \omega \mapsto \partial L(u,\omega)$ and $R : \omega \mapsto \{x \mid \inprod{z}{x} = L'(u,\omega;z)\}$.
As $Q$ is $B$-measurable~\cite[Corollary 4.6]{ROCKAFELLAR1969measurable}, we have $gr(Q)$ is $B \otimes \B(\reals^d)$-measurable by~\cite[Chapter III]{castaing2006convex}.
Moreover, $gr(R)$ is measurable as it is the graph of a linear multifunction, hence measurable.
Now we can take $gr(Q_z) = gr(Q) \cap gr(R)$, which is measurable by closure under intersections, so $gr(Q_z)$ is $B \otimes \B(\reals^d)$-measurable, and hence $Q_z$ is $B$-measurable~\cite[Chapter III]{castaing2006convex}.

\iffalse
\jessiet{If $R$ was compact}
As measurable functions are closed under (countable) intersections, we can simply show $Q_z$ intersection of measurable functions; namely, since we have $Q_z(\omega) = Q: \omega \mapsto \partial L(u,\omega) \cap R : \omega \mapsto \{x \in \reals^d : \inprod{z}{x} = L'(u,\omega;z)\}$.
\jessiet{Look at strong measurability or see if there's a result of closure under intersections}
\jessie{If $R$ is compact, then we are done by Castaing Valadier Prop iii.4}

Moreover, the function $Q$ is measurable by~\cite[Corollary 4.6]{ROCKAFELLAR1969measurable}.
Moreover, $R$ is measurable as $L'$ is measurable and $\inprod{z}{\cdot}$ is continuous in its second argument, hence Borel.

As $Q_z$ is the intersection of these two functions, we conclude $Q_z$ is measurable by~\cite[Prop iii.4]{castaing2006convex} and closed-valued, and therefore has a measurable selection $V \in \V_u$ such that $\inprod{z}{V(\omega)} = L'(u,\omega;z)$ for all $\omega \in \Omega$.
\fi

@@

3.
We want to show $\inprod{z}{v} = \E_p L'(u,Y,z)$.
This follows from linearity of expectation and construction of $V$; to see this, consider
\begin{align*}
\inprod{z}{v} &= \inprod{z}{\E_p V} \\
 &= \E_p \inprod{z}{V}
 \end{align*}
 The expectation can be pulled out of the inner product because $V$ is finite for all $\omega \in \Omega$ as it is contained in $\partial L(u,\omega)$, which is compact as $L$ is finite and convex for all $u \in \reals^d$.
 \begin{align*}
 \E_p \inprod{z}{V} &= \int \inprod{z}{V(\omega)} dp\omega \\
 &= \int L'(u,\omega; z) dp\omega \\
 &= \E_p L'(u,Y;z)
\end{align*}

@@

\bibliographystyle{ieeetr}
\bibliography{refs-selection}
